Multi-level modeling with Openflexo/\FML

### 2022-04-06 conclusions.tex
renaming of .tex files

We fulfill all the challenge requirements and we propose a tooling that
demonstrates the usability of our solution. We developed an ad hoc solution,
including models and metamodels, thanks to the flexibility provided by the
metametamodeling infrastructure offered by Openflexo. This infrastructure
helped us to overcome the limitations of the strict modeling paradigm mentioned
in Sect.~\ref{sec:introduction}, this is, the lack of support for different
forms of classifications and for the duality type/object. In that sense, our
solution seamlessly integrates the type-object pattern which allows us to
dynamically create different \emph{instances} of model elements (\eg, \emph{TaskType}, among others) and use them for typing other model elements. This, together with the \FML core support for (level-agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.

The features of \FML enable modeling flexibility which, coupled with an appropriate methodology, allow us to achieve 
the multi-level capabilities without dedicated tooling, in order to solve
problems which require it. We use virtual models to \enquote{implement} levels on demand.

>>>>

We fulfill all the challenge requirements and we propose a tool that
demonstrates the usability of our solution. We developed an ad hoc solution,
including models and metamodels, thanks to the flexibility provided by the
metametamodeling infrastructure offered by Openflexo. This infrastructure
helped us to overcome the limitations of the strict modeling paradigm mentioned
in Sect.~\ref{sec:introduction}, this is, the lack of support for different
forms of classifications and for the duality type/object. In that sense, our
solution seamlessly integrates the type-object pattern which allows us to
dynamically create different \emph{instances} of model elements (\eg,
\emph{TaskType}, among others) and use them for typing other model elements.
This, together with the \FML core support for (level agnostic) classical
linguistic instantiation, permitted us to merge linguistic and ontological
instantiation in order to provide a satisfactory solution to the Multi-Level
Process challenge.

The features of \FML enable a great modeling flexibility which, coupled with an
appropriate methodology, allow us to achieve the multi-level capabilities
without dedicated tool, in order to solve problems which require it. We use
virtual models to \enquote{implement} levels on demand.

### 2022-04-06 discussion.tex
renaming of .tex files

  Behaviors are continuously checked, ensuring the integrity of the models when
  changes occur. However, most of these behaviors have to be written by the
  users. Thus, the integrity of the contents relies on the users when they
  build the metamodels and the models. Sect.~\ref{sec:model} describes
  how the requirements are captured and how behaviors are implemented with
  \FML. For instance, \texttt{isAuthorizedActor} behavior is needed to enforce
  \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
  and \texttt{Task}. Note that in our approach, metamodels are built in an
 ad hoc way together with the models (co-construction), therefore the user
  also validates on an on-demand basis.

@@

Our toolset includes mechanisms to verify the models. First, syntactical
consistency is realized by cardinality checks and by typing. Second, the
constraints the users have written are continuously checked during model
edition and enactment. Thus, a part of the verification relies on the fact that
users add behaviors when they build the metamodels and the models. For
instance, \textbf{S6} requirement is constrained by an invariant implemented in
\texttt{CodeArtifact}. It ensures that \texttt{artifactType} has the right type
and that it implements the \texttt{languages} relation.
%\noteJC{updated}

>>>>

  Behaviors are continuously checked, ensuring the integrity of the models when
  changes occur. However, most of these behaviors have to be written by the
  users. Thus, the integrity of the contents relies on the users when they
  build the metamodels and the models. Sect.~\ref{sec:model} describes
  how the requirements are captured and how behaviors are implemented with
  \FML. For instance, \texttt{isAuthorizedActor} behavior is needed to enforce
  \reqp{17} requirement, leading to writing \FML code in \texttt{TaskType}
  and \texttt{Task}. Note that in our approach, metamodels are built in an
 ad hoc way together with the models (co-construction), therefore the user
  also validates on an on-demand basis.

@@

Our toolset includes mechanisms to verify the models. First, syntactical
consistency is realized by cardinality checks and by typing. Second, the
constraints the users have written are continuously checked during model
edition and enactment. Thus, a part of the verification relies on the fact that
users add behaviors when they build the metamodels and the models. For
instance, \reqs{6} requirement is constrained by an invariant implemented in
\texttt{CodeArtifact}. It ensures that \texttt{artifactType} has the right type
and that it implements the \texttt{languages} relation.
%\noteJC{updated}

### 2022-04-06 relatedworks.tex
renaming of .tex files

A plethora of multi-level modeling approaches and
tools\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}
with different foundations have appeared in recent
years~\parencite{somogyi2021playground}. Comparing them lies out of the scope
of this paper. In this sense, we limit the following discussion to the
presentation and comparison of previous solutions to the MULTI Process
Challenge.

>>>>

A plethora of multi-level modeling approaches and
tools\footnote{\raggedright\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}
with different foundations have appeared in recent
years~\parencite{somogyi2021playground}. Comparing them lies out of the scope
of this paper. In this sense, we limit the following discussion to the
presentation and comparison of previous solutions to the MULTI Process
Challenge.
